Automated Synthesis of State-Space Digital Systems

### 2016-12-21 cav2017/cav2017.tex
Some fixed based on conversations with Cristina

%-------------------------------
\section{Preliminaries}
%-------------------------------
In this study, we focus on Linear Dynamical Cyber-Physical Systems with
continuous plants described by Ordinary Differential Equations (ODEs) 
in the form
%
\begin{align}
\label{eq:ode}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\\
\vec{y}(t)&=\mat{C}\vec{x}+\mat{D}\vec{u},\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, 
$\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a digital system. Eq.~\ref{eq:ode} is soundly discretized 
(as shown in the appendix) into 
\textcolor{red}{Dario: Can you please describe all terms used in the equation?}
%
\begin{align}
\label{eq:plant}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k,\nonumber
\end{align}
%
and digital controllers in the form
\textcolor{red}{Dario: Can you please describe all terms used in the equation?}
%
\begin{equation}
\vec{u}_k=\mat{K}_d\vec{y}_k.
\end{equation}
%
The output of the plant is being sampled by an ADC, and since the full state of the
plant cannot be witnessed directly, the controller software includes an observer that 
estimates the state of the plant (see appendix for details on how such an observer
is constructed). We further presume that there are small tolerances in the system 
described in the following form
\textcolor{red}{Dario: Can you please describe all terms used in the equation?}
%
\begin{align*}
\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J
\mat{B} \in \mathcal{G}_B.
\end{align*}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Synthesisig the dynamics}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{A}_o,\mat{B}_o,\mat{C}_o$ by rounding down the coefficients $*_{oij}=\mathcal{F}\langle I,F,E\rangle (*_{ij}) : *=\{a,b,c\}$
\item Calculate $\mat{T}$ and $\mat{W}_o^{-1}\mat{W}_{of}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}\eqref{eq:wnof}\eqref{eq:wof}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=\mat{T}\mat{A}\mat{T}^-1$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\end{enumerate}
\item Next we synthesise a candidate controller and observer. Let $\mathcal{J}\langle I,F,E \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F,E\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F,E \rangle (P_{a-k},\mat{T})$ that satisfies Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item Select the observer error characteristic polynomial $\tilde{\mat{L}} \in \mathbb{R}\langle I,F,E\rangle^n$. 
This is achieved by solving a SAT formula $\mathcal{J}_L=\mathcal{J}\langle I,F,E \rangle (P_p,\mat{W}_o^{-1}\mat{W}_{of})$ 
that satisfies Jury's criteria for 
$P_p=z^n+\sum_{i=1}^n p_i z^{n-i} : p_i \in \vec{P}= (\mat{W}_o^{-1}\mat{W}_{of})^-1 \tilde{\mat{L}} 
+ [\begin{array}{cccc}a_1&a_2&\hdots&a_n\end{array}]^T$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\Psi_{\lambda}$ .
\item Transform the soft constraints on the eigenvalues defined in the counterexample into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \Psi_{\lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser and repeat the loop again.

@@

\begin{figure}
\centering
\hspace*{-1cm}
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.2cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=yellow!20,align=center] (pre) {{\sc pre-}\\{\sc processing}};
   \& \node[fill=yellow!20,align=center] (abstract) {\sc abstraction};
        \node[draw=none] (FIK) at ([xshift=2.5cm,yshift=.3cm]abstract)  {\sc $\Psi (\mat{K})$};
   \& \node[fill=yellow!20,align=center] (counter) {\sc counterexample};\\
   \& 
   \& \node[fill=yellow!20,align=center, minimum width=3cm] (synth) {\sc synthesiser};
       \node[draw=none] (KL) at ([xshift=2.5cm,yshift=-.25cm]synth)  {\sc $\mat{K},\mat{L}$};
       \node[draw=none] (NKL) at ([xshift=2cm,yshift=1.2cm]synth)  {\sc $\neg \mat{K}, \mat{L}, \mat{A}$};
   \& \node[fill=yellow!20,align=center, minimum width=4cm] (verify) {\sc verifier ($\Psi$)};
       \node[draw=none] (VUNS) at ([xshift=.6cm,yshift=1cm]verify)  {\sc UNSAT};
       \node[draw=none] (SAT) at ([xshift=2.5cm,yshift=-.3cm]verify)  {\sc SAT};
   \& \node[ellipse, fill=yellow!20] (done) {{\sc Done}};\\
   \&
   \& complexnode/.pic={
     \coordinate (auxsynth);
     \node[draw,rectangle,align=center] (KSAT) at ([xshift=-.8cm]auxsynth.center) {\sc SAT\\Solver};
     \node[draw,rectangle,align=center] (LSAT) at ([xshift=.8cm]auxsynth.center) {\sc SAT\\Solver};
     \node[draw=none] (SATK) at ([xshift=-.5cm,yshift=1.5cm]KSAT)  {$\mathcal{J}_K$};
     \node[draw=none] (K) at ([xshift=.4cm,yshift=1.5cm]KSAT)  {$\tilde{\mat{K}}$};
     \node[draw=none] (SATL) at ([xshift=-.5cm,yshift=1.5cm]LSAT)  {$\mathcal{J}_L$};
     \node[draw=none] (L) at ([xshift=.4cm,yshift=1.5cm]LSAT)  {$\tilde{\mat{L}}$};
    }
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAI) at ([xshift=-1.5cm]AA.center) {\sc AI\\Model};
     \node[draw,rectangle,align=center] (AAV) at ([xshift=1.5cm]AA.center) {\sc AA\\Solver};
     \node[draw=none] (AB) at ([xshift=.4cm,yshift=1.5cm]AAI.center)  {\sc $\mat{A}_t$ $\mat{B}_{tp}$ $\mat{B}_{tn}$};     
     \node[draw=none] (MA) at ([yshift=.2cm]AA.center)  {$\mathcal{A}_t$};     
     \node[draw=none] (MB) at ([yshift=-.3cm]AA.center)  {$\mathcal{B}_{tp},\mathcal{B}_{tn}$}; 
     \node[draw=none] (AB) at ([xshift=0.3cm,yshift=1.5cm]AAV.center)  {\sl $\hat{X}^\#$};  
    }\\
   \&
   \& complexnode/.pic={
     \coordinate (UNSAT);
     \coordinate (KUNSAT)  at ([xshift=-.8cm]UNSAT.center);
     \coordinate (LUNSAT) at ([xshift=.8cm]UNSAT.center);
     \node[draw=none] (UNSATB) at ([xshift=1.6cm,yshift=.3cm]UNSAT)  {\sc UNSAT};
  }
  \& \& \coordinate (FAIL);\\
  };
  \path
    (pre.east) edge (abstract.west)
    (abstract.south) edge (synth.north)
    (synth.east) edge (verify.west)
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge ([xshift=-2.1cm]synth.south) 
    ([xshift=-3.3cm]synth.south) edge ([xshift=-.6cm]KSAT.north)
    ([xshift=.6cm]LSAT.north) edge ([xshift=3.3cm]synth.south)
    ([xshift=+2.1cm]synth.south) edge ([xshift=-.6cm]LSAT.north)
    ([xshift=-4.9cm]verify.south) edge (AAI.north)
    (AAI.east) edge (AAV.west)
    (AAV.north) edge ([xshift=4.9cm]verify.south)
    (verify.north) edge (counter.south)
    (counter.west) edge (abstract.east)
    (counter.south west) edge (synth.north east)
    (FAIL.north) edge (done.south);
   \path[-] 
     (KSAT.south) edge (KUNSAT.north)
     (LSAT.south) edge (LUNSAT.north)
     (KUNSAT.east) edge (LUNSAT.west)
     (LUNSAT.east) edge (FAIL.west);
\end{tikzpicture}
\caption{The CEGIS loop with CEGAR\label{fig:CEGARIS}.}
\end{figure}

@@

\begin{table}[HTB]

>>>>

%-------------------------------
\section{Preliminaries}
%-------------------------------
In this study, we focus on Linear Dynamical Cyber-Physical Systems with
continuous plants described by Ordinary Differential Equations (ODEs) 
in the form
%
\begin{align}
\label{eq:ode}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\ \ :\ \ \vec{x} \in \mathbb{R}^{n}, \vec{u} \in \mathbb{R}^m, \mat{A} \in \mathbb{R}^{n \times n},\mat{B} \in \mathbb{R}^{n \times m}\\
\vec{y}(t)&=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t)\ \ :\ \ \vec{y} \in \mathbb{R}^{o}, \mat{C} \in \mathbb{R}^{o \times n}, \mat{D}  \in \mathbb{R}^{o \times m}\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, $\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a continuous plant. Eq.~\ref{eq:ode} is soundly discretized 
(as shown in the appendix) into
%
\begin{align}
\label{eq:plant}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k,\nonumber
\end{align}
%
and digital controllers in the form
\textcolor{red}{Dario: Can you please describe all terms used in the equation?}
%
\begin{equation}
\vec{u}_k=\mat{K}_d\vec{y}_k\ \ :\ \ \mat{K}_d \in \mathbb{R}^{m \times o}.
\end{equation}
%
The output of the plant is being sampled by an ADC, and since the full state of the
plant cannot be witnessed directly, the controller software includes an observer that 
estimates the state of the plant (see appendix for details on how such an observer
is constructed). We further presume that there are small tolerances in the system 
described in the following form
\textcolor{red}{Dario: Can you please describe all terms used in the equation?}
%
\begin{align*}
\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J \wedge
\mat{B} \in \mathcal{G}_B.
\end{align*}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Synthesisig the dynamics}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{A}_o,\mat{B}_o,\mat{C}_o$ by rounding down the coefficients $*_{oij}=\mathcal{F}\langle I,F,E\rangle (*_{ij}) : *=\{a,b,c\}$
\item Calculate $\mat{T}$ and $\mat{W}_o^{-1}\mat{W}_{of}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}\eqref{eq:wnof}\eqref{eq:wof}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Produce an initial abstraction which corresponds to an unconstrained solution space.
\end{enumerate}
\item Next we synthesise a candidate controller and observer. Let $\mathcal{J}\langle I,F,E \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F,E\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F,E \rangle (P_{a-k},\mat{T})$ that satisfies Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item Select the observer error characteristic polynomial $\tilde{\mat{L}} \in \mathbb{R}\langle I,F,E\rangle^n$. 
This is achieved by solving a SAT formula $\mathcal{J}_L=\mathcal{J}\langle I,F,E \rangle (P_p,\mat{W}_o^{-1}\mat{W}_{of})$ 
that satisfies Jury's criteria for 
$P_p=z^n+\sum_{i=1}^n p_i z^{n-i} : p_i \in \vec{P}= (\mat{W}_o^{-1}\mat{W}_{of})^-1 \tilde{\mat{L}} 
+ [\begin{array}{cccc}a_1&a_2&\hdots&a_n\end{array}]^T$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\Psi(\Lambda)$. This is a constraint in the spectrum (\emph{ie} frequency response transfer function) of the closed loop dynamics. 
\item Transform the soft constraints on the spectrum into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \Psi_{\Lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser, as well as a counterexample plant coefficients ($P_a$) and repeat the loop again. We note that the Abstraction refinement is a soft constraint on the proposed model which can increase or decrease the parameter space at any time, whereas the concrete counterexample passed to the synthesiser is monotonically decreasing. The first one improves speed of the tool whereas the second ensures convergence.

@@

\begin{figure}
\centering
\hspace*{-1cm}
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.2cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=yellow!20,align=center] (pre) {{\sc 1. pre-}\\{\sc processing}};
   \& \node[fill=yellow!20,align=center] (abstract) {\sc 4. abstraction};
        \node[draw=none] (FIK) at ([xshift=2.5cm,yshift=.3cm]abstract)  {\sc $\Psi (\mat{K})$};
   \& \node[fill=yellow!20,align=center] (counter) {\sc counterexample};\\
   \& 
   \& \node[fill=yellow!20,align=center, minimum width=3cm] (synth) {\sc 2. synthesiser};
       \node[draw=none] (KL) at ([xshift=2.5cm,yshift=-.25cm]synth)  {\sc $\mat{K},\mat{L}$};
       \node[draw=none] (NKL) at ([xshift=2cm,yshift=1cm]synth)  {\sc $P_a$};
   \& \node[fill=yellow!20,align=center, minimum width=4cm] (verify) {\sc 3. verifier ($\Psi$)};
       \node[draw=none] (VUNS) at ([xshift=.6cm,yshift=1cm]verify)  {\sc UNSAT};
       \node[draw=none] (SAT) at ([xshift=2.5cm,yshift=-.3cm]verify)  {\sc SAT};
   \& \node[ellipse, fill=yellow!20] (done) {{\sc Done}};\\
   \&
   \& complexnode/.pic={
     \coordinate (auxsynth);
     \node[draw,rectangle,align=center] (KSAT) at ([xshift=-.8cm]auxsynth.center) {\sc Prog.\\Search};
     \node[draw,rectangle,align=center] (LSAT) at ([xshift=.8cm]auxsynth.center) {\sc Prog.\\Search};
     \node[draw=none] (SATK) at ([xshift=-.5cm,yshift=1.5cm]KSAT)  {$\mathcal{J}_K$};
     \node[draw=none] (K) at ([xshift=.4cm,yshift=1.5cm]KSAT)  {$\tilde{\mat{K}}$};
     \node[draw=none] (SATL) at ([xshift=-.5cm,yshift=1.5cm]LSAT)  {$\mathcal{J}_L$};
     \node[draw=none] (L) at ([xshift=.4cm,yshift=1.5cm]LSAT)  {$\tilde{\mat{L}}$};
    }
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAI) at ([xshift=-1.5cm]AA.center) {\sc AI\\Model};
     \node[draw,rectangle,align=center] (AAV) at ([xshift=1.5cm]AA.center) {\sc AA\\Solver};
     \node[draw=none] (AB) at ([xshift=.4cm,yshift=1.5cm]AAI.center)  {\sc $\mat{A}_t$ $\mat{B}_{tp}$ $\mat{B}_{tn}$};     
     \node[draw=none] (MA) at ([yshift=.2cm]AA.center)  {$\mathcal{A}_t$};     
     \node[draw=none] (MB) at ([yshift=-.3cm]AA.center)  {$\mathcal{B}_{tp},\mathcal{B}_{tn}$}; 
     \node[draw=none] (AB) at ([xshift=0.3cm,yshift=1.5cm]AAV.center)  {\sl $\hat{X}^\#$};  
    }\\
   \&
   \& complexnode/.pic={
     \coordinate (UNSAT);
     \coordinate (KUNSAT)  at ([xshift=-.8cm]UNSAT.center);
     \coordinate (LUNSAT) at ([xshift=.8cm]UNSAT.center);
     \node[draw=none] (UNSATB) at ([xshift=1.6cm,yshift=.3cm]UNSAT)  {\sc UNSAT};
  }
  \& \& \coordinate (FAIL);\\
  };
  \path
    (pre.east) edge (abstract.west)
    (abstract.south) edge (synth.north)
    (synth.east) edge (verify.west)
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge ([xshift=-2.1cm]synth.south) 
    ([xshift=-3.3cm]synth.south) edge ([xshift=-.6cm]KSAT.north)
    ([xshift=.6cm]LSAT.north) edge ([xshift=3.3cm]synth.south)
    ([xshift=+2.1cm]synth.south) edge ([xshift=-.6cm]LSAT.north)
    ([xshift=-4.9cm]verify.south) edge (AAI.north)
    (AAI.east) edge (AAV.west)
    (AAV.north) edge ([xshift=4.9cm]verify.south)
    (verify.north) edge (counter.south)
    (counter.west) edge (abstract.east)
    (counter.south west) edge (synth.north east)
    (FAIL.north) edge (done.south);
   \path[-] 
     (KSAT.south) edge (KUNSAT.north)
     (LSAT.south) edge (LUNSAT.north)
     (KUNSAT.east) edge (LUNSAT.west)
     (LUNSAT.east) edge (FAIL.west);
\end{tikzpicture}
\caption{The CEGIS loop with CEGAR\label{fig:CEGARIS}.}
\end{figure}

@@

\begin{table}[htb]
